### 2017-04-26 paper/frag_implementation.tex
Additional implementation discussion

The implementation of this project is based on the following expectations.
First, the server must support Multipath TCP. However, beyond this, no other
modifications should be required of the server. Second, the client and detour
point are assumed to be ``active'' participants in the collective, and so they
may run whatever software is necessary to participate.

The client must run a modified kernel, which supports creating MPTCP subflows
across detours. In addition, the client must also run a daemon which discovers
and negotiates with detour points, finally reporting available ones to the
kernel. These concerns remain separate for several reasons. First, communication
with detour points requires communication over the Internet. While it is
possible to do user-level networking within the kernel, it is much simpler and
more future-proof to use the userspace socket API. Second, one type of detour
requires starting an OpenVPN process, an operation which should only be done
from userspace. Finally, userspace tools may be more easily configured and can
perform a more broad spectrum of computations, such as floating point math.

The detour need not support Multipath TCP. However, it must run one of two
possible daemons, depending on which type of detour it will be hosting.

@@

In this project, we implement two different mechanisms for tunneling MPTCP
subflows across a detour. In both cases, the detour host must run a daemon that
assists in tunneling traffic from the client to the server.

@@

\begin{itemize}
\item The connection is over UDP, to avoid the so-called ``TCP Meltdown'' effect
  caused by two congestion control algorithms interfering with each other.
  % TODO citation
\item The initial connection negotiation requires the client and server to
  exchange certificates. OpenVPN allows the user to specify the PKI root. A
  centralized detour management system could authorize clients and servers by
  signing certificates, enabling tunneling for members and preventing it for
  non-members.
\item Subsequent messages are not protected by encryption or signatures, to
  avoid the computational overhead.
\end{itemize}

@@

The OpenVPN implementation has several attractive qualities. First, it relies on
a well-used protocol for forwarding traffic. Second, it need not be configured
on a per-connection basis. Finally, it provides a built in mechanism for
mutually authenticating the client and detour. However, there are some
drawbacks. Since packets are tunneled, at least 28 bytes of overhead (IP and UDP
headers) are required. To avoid fragmentation, the connection's maximum segment
size (MSS) must be adjusted.

@@

The Linux kernel's Netfilter framework allows for two types of NAT. The first
kind is source NAT, or SNAT, rewrites the source address of the packet. This is
the typical form of NAT used by home routers: the source address of an outgoing
connection is rewritten to be the router's external IP address, but the
destination is unchanged. The second kind of NAT supported by Netfilter is
destination NAT, or DNAT. This rewrites the destination address of the packet.

@@

In order to create the tunnels, the detour runs two IPTables commands, which
create NAT rules.

>>>>

In this project, we implement a system which allows one host to dynamically add
new overlay routes to a Multipath TCP connection. Each overlay route is a new
subflow, which is tunneled across a third-party host. The system involves at
least three hosts:
\begin{itemize}
\item The \emph{client} is the host which adds detour routes, attempting to
  improve its connection quality. The client must be the host which initiates
  the MPTCP connection.
\item The \emph{server} is the host with which the client wishes to communicate.
\item The \emph{detour} is a host which acts as a waypoint along the overlay
  route from the client to the server. It tunnels the MPTCP between them. There
  may be more than one detour at a time.
\end{itemize}

We assume only that the server supports Multipath TCP, and require no additional
customization for our system to work. In our current implementation, the detour
host must run some version of Linux, and the client must run a custom version of
Linux.

The system is implemented via a number of interacting components. First, a path
manager extension for the Linux MPTCP implementation, which allows the client's
kernel to request new detour paths, receive them, and create subflows across
them. Second, a user-space utility which receives messages from the kernel and
creates the necessary detour paths. Finally, we implement two different types of
detour daemons, which run on the detour host. These components are illustrated
in Figure~\ref{f:MovingParts}.

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{figures/MovingParts.pdf}
  \caption{
    The components and their communications. In \textbf{A}, the path manager
    requests an overlay route. In \textbf{B}, the client daemon reports overlay
    routes back to the path managers. In \textbf{C} and \textbf{D}, the client
    and detour daemons negotiate an overlay route. In \textbf{E}, a subflow is
    created through the detour.
  }
  \label{f:MovingParts}
\end{figure}

The typical interaction of these components is as follows. First, the client
creates a MPTCP connection to the server. The first subflow uses the Internet
routed path. Once this connection is fully established, the path manager
requests overlay routes on which it may create subflows. The client and detour
daemons negotiate a tunnel, which may be implemented in two different ways. The
client reports this to the path-managers, which creates subflows for the MPTCP
connection across every available detour, until a predefined limit is reached.

In the remainder of this section, we will discuss each of these components at
length.

@@

We have implemented two different mechanisms for tunneling MPTCP subflows across
a detour. In both cases, the detour host must run a daemon that assists in
tunneling traffic from the client to the server.

@@

\begin{itemize}
\item The connection is over UDP, to avoid the so-called ``TCP Meltdown'' effect
  caused by two congestion control algorithms interfering with each other.
  % TODO citation
\item The initial connection negotiation requires the client and server to
  exchange certificates.
\item Subsequent messages are not protected by encryption or signatures, to
  avoid overhead.
\end{itemize}

@@

The OpenVPN implementation has several attractive qualities. First, it relies on
a well-used protocol for forwarding traffic. Second, it need not be configured
on a per-connection basis. Finally, it provides a built in mechanism for
mutually authenticating the client and detour. However, there are some
drawbacks. Since packets are tunneled, at least 28 bytes of overhead (IP and UDP
headers) are required. To avoid fragmentation, the connection's maximum segment
size (MSS) may need to be adjusted, although in our testing, this was not
required.

@@

The Linux kernel's Netfilter framework allows for two types of NAT. The first
kind, source NAT (SNAT), rewrites the source address of the packet. This is
the typical form of NAT used by home routers: the source address of an outgoing
connection is rewritten to be the router's external IP address, but the
destination is unchanged. The second kind of NAT supported by Netfilter is
destination NAT, or DNAT. This rewrites the destination address of the packet.

@@

In order to create the tunnels, the detour runs two IPTables commands, which
create NAT rules. The first performs DNAT:

\begin{lstlisting}
iptables -t nat -A PREROUTING \
         -s (*\emph{ClientAddress}*) \
         -d (*\emph{DetourAddress}*) \
         -p tcp --dport (*\emph{DetourPort}*) \
         -j DNAT --to (*\emph{RemoteAddress}*):(*\emph{RemotePort}*)
\end{lstlisting}

In this command, we add to the \texttt{nat} table, on the \texttt{PREROUTING}
chain. Incoming packets from the client, addressed to the agreed upon
\textit{DetourPort} are sent to the \texttt{DNAT} chain, which rewrites the
destination address to be the agreed upon remote address and port. This must
occur before the kernel routes the packet. Once the kernel has routed the
packet, the DNAT rule can be applied:

\begin{lstlisting}
iptables -t nat -A POSTROUTING \
         -s (*\emph{ClientAddress}*) \
         -d (*\emph{RemoteAddress}*) \
         -p tcp --dport (*\emph{RemotePort}*) \
         -j SNAT --to (*\emph{DetourAddress})
\end{lstlisting}

Since the destination address and port have already been modified, this rule
looks for packets addressed to the remote address and port, but still using the
client address as the source. It jumps these packets to the DNAT chain, where
their source address will be rewritten as the detour's address.

It is important to understand that these rules are generally only applied to the
initial SYN segment of the TCP flows. Once these rules are applied, the
connection tracking mechanism of the kernel takes over, ensuring that these
rules are applied in reverse to response packets.

The detour daemon is implemented as a Python script. It listens for tunnel
requests, executes the correct commands to create them, and then sends
responses. It ensures that detour ports do not collide with each other, or with
ports that are open on the daemon machine. It also ensures that on termination,
all tunnels are torn down.

This mechanism has the advantage that it requires zero overhead. Since no
headers are added, the full amount of data may be transmitted on these segments.
However, its disadvantage is that tunnels must be configured each time a new
connection is created on the client. While the client may simply send packets to
a new remote through an OpenVPN tunnel, it must incur a RTT delay requesting a
new tunnel from the NAT detour before it can begin establishing a subflow.

\subsection{Path Manager}

To enable subflows to use the detours created above, we implement a custom MPTCP
path manager within the Linux kernel. The path manager is a module which
dictates local policy for creating new subflows and announcing local addresses.
For example, the \texttt{fullmesh} path manager creates a subflow between each
address on the client and server.

The path manager receives space within the MPTCP control buffer in order to
track state, and it registers a set of callbacks with the protocol, so that it
can be notified of events as they occur. Our path manager (named
\texttt{detour}) runs a worker when a client-side MPTCP connection becomes fully
established (i.e. after the three-way handshake completes). The first thing this
is send a message to userspace, requesting a detour. Then, it looks through an
internal list of detours. The kernel maintains two lists of detours (per network
namespace). The first is a list of OpenVPN detours. These detour entries are not
specific to a server, and they simply indicate the name of a network interface
to use. The second list contains NAT detours. These consist of a detour address
and port, and a server address and port. The worker may only select NAT entries
which match the server address and port of the connection it is working on.

The task attempts to add up to $N$ subflows, where $N$ is a configurable limit.
It terminates in one of two ways. Either it has exhaused all of the available
detours, but not yet created all $N$ subflows. Otherwise, if it creates $N$
subflows successfully, it stops looking through the list, even if more detours
are available. If OpenVPN detours are available, then they take precedence over
the NAT detours, because they are available immediately. NAT detours created by
the client daemon become available only once the daemon has processed the
request from the kernel.

The kernel listens for new detour routes. When it receives them, it adds them to
the appropriate internal list. Then, it iterates over every MPTCP connection
within the network namespace of the message, waking their workers to give them
the opportunity to create a new subflow along this detour. If the detour was
OpenVPN type, every worker notified. If the detour was NAT type, then only
connections which match the server address and port are notified.

\subsection{Client Daemon}

The client daemon has several responsibilities. First, it must create OpenVPN
connections report them to the kernel. Second, it must respond to detour
requests from the kernel by creating NAT detours and reporting them back.

The daemon has a configuration file which allows the user to specify the IP
addresses of hosts to use for both types of detours. On startup, the daemon
connects to each OpenVPN detour and reports the resulting network interfaces to
the kernel. Then, it waits listening for detour requests from the kernel. For
each request, it requests a detour from every NAT detour address. Upon receiving
responses, it sends them the the kernel as well.

\subsubsection{Client Communication}

In order to communicate between kernel and userspace, we use Generic Netlink
sockets. Netlink is an address family for the socket interface
(\texttt{AF\_NETLINK}, similar to \texttt{AF\_INET}), which allows communication
between kernel and userspace. Many existing Linux network configuration
utilities communicate with the kernel over the Netlink address family, such as
route and firewall configuration. Each of the subsystems being configured has a
corresponding protocol family, such as \texttt{NETLINK\_NETFILTER} for Netfilter
firewall configuration. These protocol families take the same place in the
\texttt{socket()} system call as protocol families such as \texttt{IPPROTO\_TCP}
in the \texttt{AF\_INET} and \texttt{AF\_INET6} address families.

Creating a new Netlink protocol would involve allocating a new Netlink protocol
family. To make this process easier, the Generic Netlink protocol family was
created. It allows custom protocols (``Generic Netlink families'') to be
created, dynamically allocated, and discovered. It also provides a common format
for sending several data types, and creates a framework for defining message
types. Further, Generic Netlink provides a multicast capability, allowing kernel
and user sockets to broadcast messages to any socket which has joined a
particular multicast group. Finally, Generic Netlink comes with a kernel and
userspace library which allows for a consistent API and a level of abstraction
when defining Generic Netlink families.

To allow the kernel to request detours, and userspace to communicate the
existence of these detours, we define a Generic Netlink family. This protocol is
similar to the UDP protocol which the client daemon uses to request NAT tunnels.
The kernel sends a message of type \texttt{DETOUR\_C\_REQ} to a ``multicast
group'' which the client daemon subscribes to. This message contains the
parameters necessary for the daemon to create a tunnel: the server address and
port. The client daemon receives this message and creates tunnels, based on a
configuration file. It then reports these tunnels to the kernel, using a
\texttt{DETOUR\_C\_ADD} message type. The client daemon may also use the
\texttt{DETOUR\_C\_DEL} message type to delete a tunnel entry it has previously
added. Finally, the \texttt{DETOUR\_C\_ECHO} message type requests that the
kernel log all tunnels for diagnostics.
